---
title: "Priming ACE - Study 1"
output: html_document
---

############################################################
###############CLEANING###############
############################################################

Initializing packages, functions and reading in data
```{r}
library(ggplot2)
library(lavaan)

##Functions that I have created 

#Fence takes outliers and replaces at a pre-set value
fence <- function(data,var){
  myvar <- data[[var]]
  m <- median(myvar,na.rm=TRUE)
  myvar_iqr <- IQR(myvar,na.rm=TRUE)
  ul <- m+1.5*myvar_iqr
  ll <- m-1.5*myvar_iqr
  myvar[which(myvar>ul)] <- ul
  myvar[which(myvar<ll)] <- ll
  return(myvar)
}

#This function allows us to assess interaction effects by obtaining confidence intervals for
#the effects of the experiment and different levels of a moderating variable
conditional.effects <-  function(model, x, m, quantiles = 10){
  interact = paste0(x,':',m)
  beta.hat = coef(model) 
  covs = vcov(model)
 # return(covs)
  z0  = quantile(model$model[,m], seq(0 , 1, 1/quantiles))
  dy.dx = beta.hat[x] + beta.hat[interact]*z0
  se.dy.dx = sqrt(covs[x, x] + z0^2*covs[interact, interact] + 2*z0*covs[x, interact])
  upr = dy.dx+1.96*se.dy.dx
  lwr = dy.dx-1.96*se.dy.dx
  data.frame(m = z0, b = dy.dx, lwr, upr)
}

#This scales variables so we can obtain standardized effect sizes
scalerename <- function(varname,df=survey_data){
  newname <- paste0(varname,"_std")
  survey_data[[newname]] <- scale(survey_data[varname])
  survey_data
}

survey_data_old <- readRDS("C:/Users/jmoor/OneDrive/Documents/R/TOSA/tosa_survey_merged.RDS")
survey_data_old$grad <- survey_data_old$STATUS %in% c("Associate","Intern")
survey_data <- survey_data_old[which(survey_data_old$STATUS %in% c("Master Student","Student")),]
#including only those who are currently students at TOSA (not graduates)
```

Further cleaning on demographic variables; fencing and centering key variables
```{r}

#Cleaning Demographic Variables
survey_data$race_simple <- survey_data$RACE
survey_data$race_simple[which(survey_data$RACE %in% c("Native","Pacific Islander","Asian"))] <- "Other"
survey_data$white <- survey_data$race_simple=="Caucasian"
survey_data$hispanic <- survey_data$race_simple=="Hispanic"
survey_data$black <- survey_data$race_simple=="African American"
survey_data$months <- survey_data$DAYS/30
survey_data$female <- survey_data$GENDER=="Female"
survey_data$fromprison <- survey_data$ORIGIN=="Jail/Prison"
survey_data$ln_charges <- log(survey_data$X..OF.CHARGES+1)

#Fencing Outliers
fence_vars <- c("tosa_ace","gc","tosa_trust","tosa_independance","selfesteem",covs1)
for (x in fence_vars){
  survey_data[[x]] <- fence(survey_data,x)
}

#Scaling dependent variables to obtain standardized effect sizes
dvs <- c("tosa_trust","tosa_independance","selfesteem")
for (x in dvs){
  survey_data <- scalerename(x)
}

#Creating interaction terms

survey_data$aceprime <- as.numeric(survey_data$order==1) #creating intuitive variable name from conditions assignment
survey_data$ace_c <- survey_data$tosa_ace - mean(survey_data$tosa_ace,na.rm=TRUE) #centering our key moderator
survey_data$primeXace <- survey_data$aceprime*survey_data$tosa_ace
survey_data$primeXace_c <- survey_data$aceprime*survey_data$ace_c

```

############################################################
###############Preliminary Analyses###############
############################################################

Demographics
```{r}
describe(survey_data[c("months","AGE","X..OF.CHARGES")])

prop.table(table(survey_data$any_violent))
prop.table(table(survey_data$no_crime))
prop.table(table(survey_data$GENDER))
prop.table(table(survey_data$race_simple)) #86% white, 6% Black, 6% Hispanic

table(survey_data$SETS.OF.CHARGES)
```

Randomization check: Looking for whether demographic variables are balanced across conditions
```{r}
demvars <- c("any_violent","no_crime","AGE","X..OF.CHARGES","fromprison","months","female","white","hispanic","black","gc","tosa_sc")

aggregate(survey_data[demvars],list(survey_data$aceprime),FUN=mean,na.rm=TRUE)
for (x in demvars){
  myformula <- paste0(x,"~aceprime")
  print(x)
  print(summary(lm(myformula,survey_data)))
}
```


Descriptives/reliability
```{r}
cor.test(survey_data$trust_1,survey_data$trust_2)
cor.test(survey_data$indep_1,survey_data$indep_2)
psych::alpha(survey_data[c("ace_1","ace_2","ace_3r")],check.keys=TRUE)
```

############################################################
###############Primary Analyses###############
############################################################

Main DV of interest Looking at the effect of being assigned to
think about one's childhood on Trust and its interaction with self-reported adverse childhood
experiences

```{r}
mean(survey_data$tosa_ace) #2.65
sd(survey_data$tosa_ace) #.866

#this is the primary model we focus on throughout study 1
model1a <- "tosa_trust~x*aceprime+m*ace_c+xm*primeXace_c+AGE+female+ln_days+fromprison+ln_charges+white
            simpleeffect_high_ace := x+xm*.866
            simpleeffect_low_ace := x+xm*-.866" #this estimates the effect of condition at high and low levels of adverse childhood experiences (1 SD above and 1 SD below the mean)

output <- sem(model1,data=survey_data,fixed.x=FALSE,missing="FIML")#,se="bootstrap",bootstrap=300)
summary(output,standardized=TRUE)

#this model  is the exact same, but the dependent variable is made more interpretable by standardizing while we use non-cetnered adverse childhood experiences
model1b <- "tosa_trust_std~x*aceprime+m*tosa_ace+xm*primeXace+AGE+female+ln_days+fromprison+ln_charges+white
            simpleeffect_high_ace := x+xm*(2.65+.866)
            simpleeffect_low_ace := x+xm*(2.65-.866)"
output <- sem(model1,data=survey_data,fixed.x=FALSE,missing="FIML")#,se="bootstrap",bootstrap=300)
summary(output,standardized=TRUE)

#exploring the effect at different levels of adverse childhood experiences

model2 <- "tosa_trust_std~x*aceprime+m*tosa_ace+xm*primeXace+AGE+female+ln_days+fromprison+ln_charges+white
            simpleeffect_ace10 := x+xm*(1.333)
            simpleeffect_ace20 := x+xm*(2.0)
            simpleeffect_ace30 := x+xm*(2.17)
            simpleeffect_ace40 := x+xm*(2.33)
            simpleeffect_ace50 := x+xm*(2.67)
            simpleeffect_ace60 := x+xm*(3.0)
            simpleeffect_ace70 := x+xm*(3.33)
            simpleeffect_ace80 := x+xm*(3.33)
            simpleeffect_ace90 := x+xm*(3.67)"

output <- sem(model2,data=survey_data,fixed.x=FALSE,missing="FIML")#,se="bootstrap",bootstrap=300)
summary(output,standardized=TRUE)
```

Getting estimated means for figure (slightly different, more intuitive analysis)
```{r}

model1 <- "tosa_trust~x*aceprime+m*ace_c+xm*primeXace_c+AGE+female+ln_days+fromprison+ln_charges+white
            simpleeffect_high_ace := x+xm*.866
            simpleeffect_low_ace := x+xm*-.866"

#model without covariates (better captures the relationship between adverse childhood experiences and the dependent variable, makes for clearner and more intuiritive figure)
model2 <- "tosa_trust~x*aceprime+m*ace_c+xm*primeXace_c
            simpleeffect_high_ace := x+xm*.866
            simpleeffect_low_ace := x+xm*-.866"

output <- sem(model1,data=survey_data,fixed.x=FALSE,missing="FIML")#,se="bootstrap",bootstrap=300)
summary(output,standardized=TRUE)


mean(survey_data$tosa_trust,na.rm=TRUE) #4.03 is the mean overall

#intercept: 4.14
#ace prime : -.2
#ace: .14
#int: -.48

#High ACE
4.14+.14 #4.28
4.14+.14-.2-.48 #3.6

#Low ACE
4.14-.14 #4
4.14-.14-.2+.48 #4.28

sd(survey_data[which(survey_data$aceprime==TRUE),"tosa_trust"],na.rm=TRUE)
```


Looking at effects on the secondary dependent variable
```{r}
#non-standardized
model2 <- "tosa_independance~x*aceprime+m*tosa_ace+xm*primeXace+AGE+female+ln_days+fromprison+ln_charges+white
            simpleeffect_high_ace := x+xm*+.866
            simpleeffect_low_ace := x+xm*-.866"

#standardized
model2 <- "tosa_independance_std~x*aceprime+m*tosa_ace_c+xm*primeXace_c+AGE+female+ln_days+fromprison+ln_charges+white
            simpleeffect_high_ace := x+xm*(2.65+.866)
            simpleeffect_low_ace := x+xm*(2.65-.866)"

output <- sem(model2,data=survey_data,fixed.x=FALSE,missing="FIML")#,se="bootstrap",bootstrap=300)
summary(output,standardized=TRUE)

#exploring the effect at different levels of adverse childhood experiences
model2 <- "tosa_independance_std~x*aceprime+m*tosa_ace+xm*primeXace+AGE+female+ln_days+fromprison+ln_charges+white
            simpleeffect_ace10 := x+xm*(1.333)
            simpleeffect_ace20 := x+xm*(2.0)
            simpleeffect_ace30 := x+xm*(2.17)
            simpleeffect_ace40 := x+xm*(2.33)
            simpleeffect_ace50 := x+xm*(2.67)
            simpleeffect_ace60 := x+xm*(3.0)
            simpleeffect_ace70 := x+xm*(3.33)
            simpleeffect_ace80 := x+xm*(3.33)
            simpleeffect_ace90 := x+xm*(3.67)"

output <- sem(model2,data=survey_data,fixed.x=FALSE,missing="FIML")#,se="bootstrap",bootstrap=300)
summary(output,standardized=TRUE)

quantile(survey_data$tosa_ace,c(.1,.2,.3,.4,.5,.6,.7,.8,.9))

```



###############################################################################################
################Extra/Exploratory Analyses not used in Manuscript######################
###############################################################################################
```{r}
aceint_trust <- lm(tosa_trust_std~aceprime*tosa_ace+AGE+female+ln_days+fromprison+white,survey_data)

con_effects <- conditional.effects(aceint_trust,x="aceprime",m="tosa_ace",quantiles=10)
test <- con_effects[2:10,] #this was with 10 quantiles
test2 <- con_effects[4:17,]

test$m <- c(.1,.2,.3,.4,.5,.6,.7,.8,.9)


(mod_plot <- ggplot(test[2:8,], 
       aes(x = m, 
           y = b, 
           ymin = lwr,
           ymax = upr,
           xmin = .2,
           xmax=.8)) +
  geom_smooth(stat = "identity",
              color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  theme_bw() +
  labs(x = "Childhood Mistreatment",
       y = "Priming Effect on General Trust"))

```

Effects on Self-esteem (not the primary focus of the analysis, this one was just exploratory)
```{r}

summary(lm(selfesteem~(order==1),survey_data))
summary(lm(selfesteem~(order==1)*scale(tosa_ace),survey_data))


```

Looking at interaction with specific trust (this is something that was interesting but ultimately was left off the manuscript)
```{r}
summary(lm(tosa_trust_std~aceprime*gc,survey_data))

summary(lm(tosa_trust_std~aceprime*tosa_gc_1,survey_data))
summary(lm(tosa_trust_std~aceprime*tosa_gc_2,survey_data))

summary(lm(tosa_independance~aceprime*gc_5,survey_data))


survey_data$gc2Xprime <- scale(survey_data$tosa_gc_2)*survey_data$aceprime

model2 <- "tosa_trust~x*aceprime+m*tosa_ace+xm*gc2Xprime+tosa_gc_2+AGE+female+ln_days+fromprison+ln_charges+white
            simpleeffect_high_ace := x+xm*(2.65+.866)
            simpleeffect_low_ace := x+xm*(2.65-.866)"


summary(sem(model2,survey_data))


model2 <- "tosa_independance~x*aceprime+xm*gc2Xprime+tosa_gc_2+tosa_ace+AGE+female+fromprison+white+ln_days"


summary(sem(model2,survey_data))
        
       # +AGE+female+ln_days+fromprison+ln_charges+white))
```

